---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Riccardo Fusaroli"
date: "August 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the third exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time, then learning how to cross-validate models and finally how to systematically compare models.

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/ra99bdvm6fzay3g/demo_test.csv?dl=1
* Utterance Length data: https://www.dropbox.com/s/uxtqqzl18nwxowq/LU_test.csv?dl=1
* Word data: https://www.dropbox.com/s/1ces4hv8kh0stov/token_test.csv?dl=1

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data () and on the test data. Report both of them. Compare them. Discuss why they are different.

-Arecreate the models you chose last time (just write the model code again and apply it to your training data (from the first assignment))
B calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())
C create the test dataset (apply the code from assignment 1 part 1 to clean up the 3 test datasets)
D test the performance of the models on the test data (Tips: google the functions "predict()")
F optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())
```{r}
library(lmerTest)
library(Metrics)
library(stringr)
library(dplyr)

library(data.table)
library(ggplot2)
library(stringi)
library(stringr)
library(tidyverse)
library(zoo)
library(plyr)
library(MuMIn)
library(modelr)
library(ModelMetrics)
```


recreate the models you chose last time (just write the model code again and apply it to your training data (from the first assignment))
```{r}

merged_data <- read.csv("merged_data.csv")
model <- lmer(CHI_MLU~VISIT* Diagnosis+I(VISIT^2)*Diagnosis +I(VISIT^3)* Diagnosis +(1+VISIT+I(VISIT^2) +I(VISIT^3)|SUBJ),merged_data, REML = F)
summary(model)

#doing B
##root mean square error = rmse

preds <- predict(model, type= "response")

rmse(merged_data$CHI_MLU, preds)#0.2570438

#doiong C
demo_test<- read.csv("demo_test.csv")
#we see dots in the end of the names, extra variables, rename collumn, mullenrae into nonverbalIQ, expressivelangraw into verbalIQ
demo_test <- dplyr::rename(demo_test, SUBJ = Child.ID)
demo_test$SUBJ <- str_replace_all(demo_test$SUBJ, "\\.", "")
demo_test <- dplyr::rename(demo_test, VISIT = Visit)
demo_test <- dplyr::select(demo_test, SUBJ, VISIT, Ethnicity, Diagnosis, Gender, Age, ADOS,  MullenRaw, ExpressiveLangRaw )

demo_test <- dplyr::rename(demo_test, nonverbalIQ = MullenRaw)
demo_test <- dplyr::rename(demo_test, verbalIQ = ExpressiveLangRaw)

LU_test <- read.csv("LU_test.csv")
LU_test$SUBJ <- str_replace_all(LU_test$SUBJ, "\\.", "")
LU_test$VISIT <- str_extract(LU_test$VISIT,"[^vVisit.]")
LU_test <- dplyr::select(LU_test, SUBJ, VISIT, MOT_MLU, MOT_LUstd, CHI_MLU, CHI_LUstd)

token_test<- read.csv("token_test.csv")
token_test$VISIT <- str_extract(token_test$VISIT,"[^vVisit.]")
token_test$SUBJ <- str_replace_all(token_test$SUBJ, "\\.", "")
LU_test$VISIT <- as.numeric(LU_test$VISIT)
token_test$VISIT <- as.numeric(token_test$VISIT)

m1 <- full_join(LU_test, token_test, by = c("SUBJ", "VISIT"))

demo_test_real <- left_join(m1, demo_test, by = c("SUBJ", "VISIT"))
demo_test_real$X <- NULL





```
calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())

```{r}

##root mean square error = rmse
preds <- predict(model, type= "response")

rmse(merged_data$CHI_MLU, preds)#0.2593025
#this is a raw measure of the size of the error of your prediction - everytime you make a prediction you expect it to be off by 0.2593025 (the smaller the better)
#test data we have random effect for kids individually and for train we don't
#why is it bigger number for train data? we look for specifity for kids
#simple model was worse it has higher error

```

test the performance of the models on the test data (Tips: google the functions "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())
```{r}
##root mean square error = rmse
preds1 <- predict(model, newdata=demo_test_real,allow.new.levels=T)

rmse(demo_test_real$CHI_MLU, preds1) #0.7222786

      
```


formatting tip: If you write code in this document and plan to hand it in, remember to put include=FALSE in the code chunks before handing in.

[HERE GOES YOUR ANSWER]

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

#how many bottles of wine we want to buy every week? 
#we have 3 predictors how many bottles have been bought in the previous week etc...
#we fit the 3 models with three predictors which one is the best?
#are we overfitting? we can not split data from previous week and we want to use all the data because we can not split we do not have enought data to test and train
#cross validation - test and train (half)  pizza example
cross validation - allows us to estimate test error
#we have 3 models and we get where is the lowest error 


#how do we do that?
pizza - 5 slices - each kid in the slice- we want to put int he slces unique child id -createfold(data$child_id, k = (number of fold with unique kids))
-we want a loop process to look through each single slice
for (f in fold ) -> we want to have train (data$child.id is not in f -> subset(data, !(child.id %in% = F)) and test data(the same but without !)

then you create your model lmer on your train data then rmse train, rmse test
set a counter M= 1 and before we close the loop add one to the counter 
error - we want to create rmsetrain = NULL and the same for test


In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Use cross-validation to compare your model from last week with the basic model (Child MLU as a function of Time and Diagnosis, and don't forget the random effects!)
- (Tips): google the function "createFolds";  loop through each fold, train both models on the other folds and test them on the fold)

```{r}

#merging data 
demo_test_real$X <- NULL
demo_test_real$types_shared <- NULL
merged_data$X <- NULL

full_data = rbind(demo_test_real, merged_data)

write.csv(full_data, file = "full_data.csv")

full_data$SUBJ = as.factor(full_data$SUBJ)
full_data$SUBJ = as.integer(full_data$SUBJ)


#Create fold with unique SUBJ
library(caret)
folds=createFolds(full_data$SUBJ, k = 5)

#Create matarix to store results
cv_empty=matrix(ncol=3, nrow=5)
cv_empty1= matrix(ncol=3,nrow = 5)


#Loop
#Indexing, reset when using the loop again
n=1

set.seed(101)

for (k in folds){
#------ Split into training and test data ------ 
#Create training dataset, data not in fold k
data_train=subset(full_data, !(SUBJ %in% k))
#Create test dataset, data in fold k
data_test=subset(full_data, SUBJ %in% k)
  
#------ train model - apply model to data_train ------
model2 = lmer(CHI_MLU ~ VISIT * Diagnosis + (1+ VISIT|SUBJ), data_train)
#Make predictions based on model
Prediction=predict(model2, type='response')
#Prints the prediction against the test data
results_train=ModelMetrics::rmse(data_train$CHI_MLU, Prediction)  

#------ test the model - test model on data_test (last quarter) ------
#Make predictions based on modeVIS
Prediction_test=predict(model2, data_test, type='response', allow.new.levels = TRUE)
#
results_test=ModelMetrics::rmse(data_test$CHI_MLU, Prediction_test)
  
#------ save the performance ------ 
#Fills out the kth row in cv_empty with iteration and rmse's
cv_empty[n,]=cbind(n,results_train,results_test)

######################################################
  
#------ train second model - apply model to data_train ------
model3 = lmer(CHI_MLU~VISIT* Diagnosis+I(VISIT^2)*Diagnosis +I(VISIT^3)* Diagnosis +(1+VISIT+I(VISIT^2) +I(VISIT^3)|SUBJ),data_train)
#Make predictions based on model
Prediction1=predict(model3, type='response')
#Prints the prediction against the test data
results_train1=ModelMetrics::rmse(data_train$CHI_MLU, Prediction1)  

#------ test the model
Prediction_test1=predict(model3, data_test, type='response', allow.new.levels = TRUE)
#
results_test1=ModelMetrics::rmse(data_test$CHI_MLU, Prediction_test1)
  
#------ save the performance ------ 
cv_empty1[n,]=cbind(n,results_train1,results_test1)
n=n+1
}



#------ Make the matrix into a data frame ------ 
cv_results=data.frame(cv_empty)
cv_results1=data.frame(cv_empty1)
#Rename variables 
cv_results=setnames(cv_results,"X1", "iteration")
cv_results=setnames(cv_results,"X2", "train")
cv_results=setnames(cv_results,"X3", "test")
cv_results1=setnames(cv_results1,"X1", "iteration1")
cv_results1=setnames(cv_results1,"X2", "train1")
cv_results1=setnames(cv_results1,"X3", "test1")

results_models <- data.frame(mean(cv_results$train), mean(cv_results$test), mean(cv_results1$train), mean(cv_results1$test))

results_models=setnames(results_models,"mean.cv_results.train.", "train_1st")
results_models=setnames(results_models,"mean.cv_results.test.", "test_1st")
results_models=setnames(results_models,"mean.cv_results1.test.", "test_2nd")
results_models=setnames(results_models,"mean.cv_results1.train.", "train_2nd")

#after calculating the mean rmse for cross validation, we can see that the second model performed a little bit better, even though the difference is minimal


```



pizza - 5 slices - each kid in the slice- we want to put int he slces unique child id -createfold(data$child_id, k = (number of fold with unique kids))
-we want a loop process to look through each single slice
for (f in fold ) -> we want to have train (data$child.id is not in f -> subset(data, !(child.id %in% = F)) and test data(the same but without !)

Which model is better at predicting new data: the one you selected last week or the one chosen via cross-validation this week?

- Test both of them on the test data.
- Report the results and comment on them.

- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test data


[HERE GOES YOUR ANSWER]

### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.
(Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; input the average of the TD group  for each parameter in the model as X1, X2, etc.).

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```{r}
bernie <- dplyr::filter(full_data, CHI_MLU == 1.9844560 & MOT_MLU == 3.414397)
#bernie is a subject n 14

#Creating a dataset without bernie (Bernie is number 63)
no_bernie_data = subset(full_data, (SUBJ != 14))
#ONLY bernie
only_bernie_data = subset(full_data, (SUBJ == 14))

#----- Create model and make average TD child at each visit ------
# creating final model
model_full = lmer(CHI_MLU~VISIT* Diagnosis+I(VISIT^2)*Diagnosis +I(VISIT^3)* Diagnosis +(1+VISIT+I(VISIT^2) +I(VISIT^3)|SUBJ),no_bernie_data)

#Get betas
summary(model_full)

# getting means of variables for the TD kids, used for input as "x" in model
visit_mean <- full_data  %>% filter(Diagnosis == "B") %>% dplyr::group_by(VISIT)  %>% dplyr::summarise(mean(CHI_MLU))

MLU_visit1 = 0.72 + 0.71*1.31 - 0.17*1.31 + 0.01*1.31 -0.72*1.31 +0.37*1.31   -0.04*1.31

MLU_visit2 = 0.72 + 0.71*1.76 - 0.17*1.76 + 0.01*1.76- 0.72*1.76 +0.37*1.76   -0.04 *1.76

MLU_visit3 = 0.72 + 0.71*2.23 - 0.17*2.23 + 0.01*2.23 -0.72*2.23  +0.37*2.23  -0.04*2.23

MLU_visit4 = 0.72 + 0.71*2.73 - 0.17*2.73 + 0.01*2.73 -0.72*2.73 +0.37*2.73 -0.04*2.73 

MLU_visit5 = 0.72 + 0.71*3 - 0.17*3 + 0.01*3 -0.72*3   +0.37*3 - 0.04*3 

MLU_visit6 = 0.72 + 0.71*2.91 - 0.17*2.91 + 0.01*2.91 - 0.72*2.91   +0.37*2.91  -0.04*2.91

# making a list of MLU of TD kids at the different visits as calculated above.
MLU_list = c(MLU_visit1, MLU_visit2, MLU_visit3, MLU_visit4, MLU_visit5, MLU_visit6)

#getting the difference in MLU between Bernie and the TD per visit
only_bernie_data[,5] - MLU_list #Bernie has higher MLU than average TD child


#getting the difference between Bernie and his predicted score
Prediction3=predict(model_full, only_bernie_data[6,], allow.new.levels = TRUE)

only_bernie_data[6,5]-Prediction3 #he performed better than expected


```


[HERE GOES YOUR ANSWER]

### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).

Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.

- Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?

### OPTIONAL: Exercise 5): Using Lasso for model selection

Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!

